---
title: "K-Means Clustering"
author: "Michael Moran"
date: "November 7, 2017"
output: html_document
---

# Exercise 0: Install these packages if you don't have them already

```{r}
#install.packages(c("cluster", "rattle", "rattle.data", "NbClust", "flexclust"))
```

# Now load the data and look at the first few rows

```{r}
data(wine, package="rattle.data")
head(wine)
```


# Exercise 1: Remove the first column from the data and scale it using the scale() function

```{r}
df <- scale(wine[-1])         
```

# Method 1: A plot of the total within-groups sums of squares against the 
# number of clusters in a K-means solution can be helpful. A bend in the 
# graph can suggest the appropriate number of clusters. 

```{r}
wssplot <- function(data, nc=15, seed=1234){
	              wss <- (nrow(data)-1)*sum(apply(data,2,var))
               	      for (i in 2:nc){
		        set.seed(seed)
	                wss[i] <- sum(kmeans(data, centers=i)$withinss)}
	                
		      plot(1:nc, wss, type="b", xlab="Number of Clusters",
	                        ylab="Within groups sum of squares")
	   }

wssplot(df)
```
# Exercise 2:
#   * How many clusters does this method suggest?
#   * Why does this method work? What's the intuition behind it?
#   * Look at the code for wssplot() and figure out how it works

#   This method suggests 3 clusters. The method works by comparing the within-groups sums of squares against the number of clusters and identifying where additional custers cause only a small decrease in the within-groups sums of squares. The within-groups sums of squares is the variation due to differences within individual samples.  


# Method 2: Use the NbClust library, which runs many experiments
# and gives a distribution of potential number of clusters.

```{r}
library(NbClust)
set.seed(1234)
nc <- NbClust(df, min.nc=2, max.nc=15, method="kmeans")
barplot(table(nc$Best.n[1,]),
	          xlab="Numer of Clusters", ylab="Number of Criteria",
		            main="Number of Clusters Chosen by 26 Criteria")
```


# Exercise 3: How many clusters does this method suggest? 

# This method also suggests three clusters.

# Exercise 4: Once you've picked the number of clusters, run k-means 
# using this number of clusters. Output the result of calling kmeans()
# into a variable fit.km

```{r}
fit.km <- kmeans(df, 3, nstart=25)
```

# Exercise 5: using the table() function, show how the clusters in fit.km$clusters
# compares to the actual wine types in wine$Type. Would you consider this a good
# clustering?

```{r}
ct.km <- table(wine$Type, fit.km$cluster)
ct.km
```
```{r}
library(flexclust)
randIndex(ct.km)
```


#This would be considered good clustering as the agreement is 0.90. 

# Exercise 6:
# * Visualize these clusters using  function clusplot() from the cluster library
# * Would you consider this a good clustering?

```{r}
library(cluster)
#clusplot(ct.km)
```
